---
title: "MADPH_Rx"
author: "Charlotte Chase"
date: "7/7/2020"
output: html_document
---




```{r}
library(tidyverse)
library(devtools)
library(urbnmapr)
library(readxl)
library(viridis)
library(fastDummies)
```

#reading in quarter information for 2016 to 2019

```{r}
# read in 2016-2019 data
Rx <- read_excel(
  "MADPH_PMP.xlsx",
  range = cell_cols("A:H") #cell_cols selects all rows associated w those columns
)
```

#reading in quarter 1, 2015

```{r}
# read in 1st quarter of 2015
Rx.q1 <- read_excel(
  "pmp-county-data-roll-q1-2015.xlsx",
  range = "A5:F20"
)
```

#adding in "year" and "quarter" to "Rx.q1"

```{r}
Rx.q1 <- Rx.q1 %>% 
  mutate(Year = rep(2015, nrow(Rx.q1)))  %>% #creating a new column,"Year". For every row in the column "Year" in "Rx.q1," write "2015." 
  mutate(Quarter = rep(1, nrow(Rx.q1))) 
```



#reading in quarter 2, 2015

```{r}
# read in 2nd quarter of 2015
Rx.q2 <- read_excel(
  "pmp-county-data-roll-q2-2015.xlsx",
  range = "A5:F20"
)
```

#adding in "year" and "quarter" to "Rx.q2"

```{r}
Rx.q2 <- Rx.q2 %>% 
  mutate(Year = rep(2015, nrow(Rx.q2)))  %>% #creating a new column,"Year". For every row in the column "Year" in "Rx.q2," write "2015." 
  mutate(Quarter = rep(2, nrow(Rx.q2))) 
```


#reading quarter 3, 2015

```{r}
# read in 3rd quarter of 2015
Rx.q3 <- read_excel(
  "pmp-county-data-roll-q3-2015.xlsx",
  range = "A5:F20"
)
```


```{r}
Rx.q3 <- Rx.q3 %>% 
  mutate(Year = rep(2015, nrow(Rx.q3)))  %>% #creating a new column,"Year". For every row in the column "Year" in "Rx.q3," write "2015." 
  mutate(Quarter = rep(3, nrow(Rx.q3))) 
```

#merge all of the county data together from 2015, and then to the 2016-2019

```{r}
full.rx <- Rx.q1 %>% 
  full_join(Rx.q2) %>% 
  full_join(Rx.q3) %>% 
  full_join(Rx) %>% 
  set_names( c("County", "Population", "Total Rx", "Total Rx Units", "N.People w/ Rx", "Percent of County Pop w/ Rx", "Year", "Quarter")) %>%  #renaming all of the names
  mutate(county_name = paste(County, "County")) %>%   #mutating a new column, county_name...on the other side is the expression to make a row. paste from the currently existing column, "County," and then add the string "County". It's evaluated on a row by row basis.  
  dummy_rows(select_columns = c("Year", "Quarter", "County"))
  
```


#checking the column names in "full.rx."
```{r}
colnames(full.rx)
```


#Make maps...

#Making a generic US map.
```{r}
ggplot() + 
  geom_polygon(data = urbnmapr::states, mapping = aes(x = long, y = lat, group = group),
		           fill = "grey", color = "white") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45)
```

```{r}
household_data <- left_join(countydata, counties, by = "county_fips") #joining the dataframes countydata and counties by "county_fips"

household_data %>%
  ggplot(aes(long, lat, group = group, fill = medhhincome)) +
  geom_polygon(color = NA) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Median Household Income")
```

#look at counties dataset
```{r}
str(counties)
```

#Changing California example to Massachusetts

#the data from "urbnmapr" has differences in how they name counties
```{r}
test <- c("#f2f0f7", "#dadaeb", "#bcbddc", "#9e9ac8", "#756bb1", "#54278f")

full.rx %>%
  filter("county_name" != "MA County")%>% 
  left_join(counties, by = "county_name") %>% 
  filter(state_name == "Massachusetts") %>% 
  filter(Year == 2018, Quarter == 1) %>% 
  ggplot(mapping = aes(long, lat, group = group, fill = `Percent of County Pop w/ Rx`)) +
  labs(title = "Percent of Population Prescribed Schedule II Medications, by County")+
  coord_map(projection = "albers", lat0 = 39, lat1 = 45)+
  geom_polygon(color = "#ffffff", size = .25) +
  theme_void() +
  scale_fill_viridis(colors = test) 
```


#Now, on to making the second map: deviations from the state average.


#Make a new column, showing the linear difference between the state average and the county


#"ma.percents" just has the year, the quarter, and the % of the population reciving a scheduled Rx for all of MA.

#a left join ... keeps all of the rows from the "left hand" and all the columns from both dataframes. If there are multiple matches between the left and right dataframes, all combinations are returned

#B/c "ma.percents" has the only "year" and "quarter," it's going t create a duplicate of the "ma.percent" column for each row that matches in the left hand dataframe.  

#bc each county has the same year and quarter, it creates a duplicate MA percent for each county. This is bc we did a left join. 
```{r}
ma.percents <- full.rx %>%  #making a dataframe called ma.percents
  filter(County == "MA") %>%  #we only want to look at "MA" in the County data
  select(Year, Quarter, `Percent of County Pop w/ Rx`) %>%  #select picks only "year", "quarter", and that percent
  rename(ma.percent = `Percent of County Pop w/ Rx`) #renaming so that a later join will result in a new column


full.rx %>% #piping in full.rx, and doing a left join with the dataframe that we just made 
  left_join(ma.percents) %>% #bc full.rx is piped in, it's the left hand dataframe
  mutate(difference = `Percent of County Pop w/ Rx` - ma.percent) %>% #making a column called "difference"...diff btwn county %age and MA %age
  # view()
  filter(County != "MA") %>% #removing MA county
  left_join(counties, by="county_name") %>% #joining it into the counties dataframe, which we need to make a map
  filter(state_name == "Massachusetts") %>% 
  filter(Year == 2016, Quarter == 1) %>% 
  ggplot() +
    aes(long, lat, group = group, fill = `difference`) +
    geom_polygon(color = "gray", size = 0.25) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
    theme_void() +
    scale_fill_distiller(type = "div", palette = "PRGn", na.value = "orange") + #type div = divergent--we want it to diverge from 0
    labs(title = "Difference from State % of Pop. Receiving Schedule 2 Rx by County")
```
#green = low, purple = high



#Now, shiny time.



